---
title: "Machine Learning Project"
author: "ME"
date: "4/25/2020"
output: html_document
---

```{r echo=TRUE}
library(caret)
library(corrplot)
set.seed(123)

```

## Getting Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


```{r echo=TRUE}
Training<-read.csv("pml-training.csv", stringsAsFactors = FALSE)
Testing<-read.csv("pml-testing.csv", stringsAsFactors = FALSE)
```

## Looking at the data

There's a lot of useless columns.

```{r, echo=TRUE}
dim(Training)

```

##Clean the data
Getting rid of factors that are near zero variance and have more than 50% null values.
Then getting rid of first two columns that's just row number and subject names.
I am also changing the format of the timestamp. 
```{r, echo=TRUE}
remove_col<-nearZeroVar(Training,names=TRUE)
all_col<-names(Training)
TrainingClean<-Training[setdiff(all_col,remove_col)]

##Removing factors with more than 50% null value
TrainingClean<- TrainingClean[,!sapply(TrainingClean, function(x) mean(is.na(x)))>0.5]

##Changing format 
TrainingClean$classe<-as.factor(TrainingClean$classe)

##getting rid of timestamps, row number, subject names(not helpful for prediction)
TrainingClean<- subset(TrainingClean,select = -c(1:5))
str(TrainingClean)

##checking to see whether there are NAs
names(which(sapply(TrainingClean,anyNA)))




```
##Correlations 
There are lots of variables that are correlated with each other so I will use PCA to reduce the number of factors without compromising the accuracy significantly.

```{r,echo=TRUE}
M<-abs(cor(TrainingClean[,-54]))
diag(M)<-0
which(M>0.8,arr.ind=T)

```

##Using random forest with PCA
Wih an accuracy of 98% and OOB estimate of error rate at 1.74%
```{r, echo=TRUE}

fit<-train(classe ~., method ="rf",preProcess=c("center","scale","pca"),data=TrainingClean, trControl = trainControl(method = "cv"), ntree=150)
print(fit$finalModel)
confusionMatrix(fit)
plot(fit)
testPC<-predict(fit,Testing)

```
